---
title: "Codebook"
author: "Charin Polpanumas"
date: "February 18, 2559 BE"
output:
  html_document:
    toc: true
    number_sections: true
    theme: spacelab
---

# Data Overview
The ```tidy Thai stock dataset``` includes annual fundamental data of 463 Thai stocks in ```udf.csv``` and annual return of SET TRI in ```set_return.csv``` from 2002-2015. Monetary unit is THB.

This codebook demonstrates how fundamental data on Thai stocks from the [Morningstar](http://www.morningstar.com) and SET TRI from [Stock Exchange of Thailand](http://set.or.th) are transformed into the tidy format used in this [analysis](index.html).

# Getting and Cleaning Data
## Thai Stock Data
The data on Thai stock fundamentals are obtained from [Morningstar](http://www.morningstar.com). The symbols associated wit the stock market was obtained by using a filter on the website and saved to ```symbols.csv```. We download JSON associated with this and save them to the ```stocks``` folder.
```{r, eval=FALSE}
library(jsonlite)
library(httr)

#Run script on 14/2/2016

#Token: financials, price
api_string <-'MY_API_STRING'

#Read symbols
symbols<-read.csv('symbols.csv')

#Loop writing to stocks/
for (i in symbols$x) {
    request_string<-paste0(api_string,i,'/financials')
    j<-fromJSON(request_string)
    s <-toJSON(j)
    write(s,paste0('stocks/',i))
}
```

## SET TRI and Annual Returns

The data on SET TRI (Total Return Index) are obtained from [Stock Exchange of Thailand](http://set.or.th) as ```set.html```. We seletively pick the first and last days of 2002-2015 in order to calculate the annual returns (%) from the respective years. The result is then saved to ```set_return.csv```.

```{r, eval=FALSE}
library(XML)

#Download and save as set.html
download.file(url='http://www.set.or.th/set/fetchfile.do?filename=roi/market.xls',destfile = 'set.html')

#Read as html table
se<-readHTMLTable('set.html',header=TRUE)

#Convert to data.frame
se<-as.data.frame(se)

#Edit column names
colnames(se)<-gsub('NULL.','',colnames(se))

#Convert date
se$datetime<-as.Date(se$Date,format='%d %b %Y')

#Get first and last days of 2002-2015
fdf<-data.frame()
ldf<-data.frame()
for (i in 2002:2015) {
    first<-min(se[year(se$datetime)==i,]$datetime)
    last<-max(se[year(se$datetime)==i,]$datetime)
    fdf<-rbind(fdf,se[se$datetime==as.Date(first),])
    ldf<-rbind(ldf,se[se$datetime==as.Date(last),])
}

#Calculate return for each year
f<-as.numeric(gsub(',','',levels(fdf$SET.TRI)[fdf$SET.TRI]))
l<-as.numeric(gsub(',','',levels(ldf$SET.TRI)[ldf$SET.TRI]))
final<-data.frame(i=2002:2015,set_return=(l/f-1)*100)

#Save to set_return.csv
write.csv(final,file='set_return.csv',row.names = FALSE)
```

## Cleaning and Making the Tidy Thai Stock Dataset
We process the data from ```stocks``` folder and ```set_return.csv``` to make the tidy Thai stock dataset saved in ```udf.csv```. The process includes:

* Remove rows with trailing twelve months (TTM) data.
* Convert year-month format of year indicator (```i```) to year only.
* Change variable names to lowercase, remove non-alphanumeric characters, trim spaces.
* Convert all columns except dates from character to numeric.
* Add forward price (price next year) as ```fwd_price```.
* Add price growth as ```priceg``` calculated by ```fwd_price/monthendstockprice```. Note that this price growth is forward-looking, which suits our analysis of using current year fundamentals to forecast future price growth. For other applications, use with care.
* Remove infinities and NaN, mostly resulted from creating ```priceg```.
* Aggregate data for all stocks.
* Join with ```set_return``` on year indicator(```i```) to give SET TRI growth rate as benchmark.
* Save to ```udf.csv```.


```{r, eval=FALSE}
library(jsonlite)
library(lubridate)
library(stringr)
library(plyr)
library(zoo)

#Set up list to contain each individual stock
lst<-list()

#Loop cleaning operations through all symbols
for (stock_name in dir('stocks')){
    
    #Load stock info from json
    j<-fromJSON(paste0('stocks/',stock_name))
    
    #Set symbol and year to data.frame
    lst[[stock_name]]<-data.frame(symbol=rep(stock_name,length(j$financials$annuals$`Fiscal Year`)))
    
    #Bind all elements to data.frame
    for (i in j$financials$annuals) {
        lst[[stock_name]]<-cbind(lst[[stock_name]],i)
    }
    
    #Remove TTM
    lst[[stock_name]]<-lst[[stock_name]][lst[[stock_name]][['i']]!='TTM',]
    
    #Convert yearmon to year
    lst[[stock_name]][['i']]<-year(as.yearmon(lst[[stock_name]][['i']]))
    
    #Change header to lowercase, remove non-alphanumeric characters, trim spaces
    colnames(lst[[stock_name]])<-str_to_lower(colnames(lst[[stock_name]]))
    colnames(lst[[stock_name]])<-gsub('[^a-zA-Z0-9]','',colnames(lst[[stock_name]]))
    colnames(lst[[stock_name]])<-trimws(colnames(lst[[stock_name]]))
    
    #Convert all columns except dates to numeric
    for (i in 1:length(lst[[stock_name]])){
        if (!grepl('(filing|symbol|^i$)',colnames(lst[[stock_name]])[i],ignore.case = TRUE)) {
            lst[[stock_name]][,i]<-as.numeric(as.character(lst[[stock_name]][,i]))
        }
    }
    
    #Add forward price and price growth in %
    end_price <-lst[[stock_name]][['monthendstockprice']]
    fwd_price<-c()
    for (k in 1:(length(end_price)-1)) {
        fwd_price[k]<-end_price[k+1]
    }
    fwd_price[length(end_price)]<-NA
    lst[[stock_name]]$fwd_price <- fwd_price
    lst[[stock_name]]$priceg<- round((fwd_price/end_price-1)*100,3)
    
    #Remove infinities and NaN
    lst[[stock_name]]$priceg<-gsub('(NaN|Inf|-Inf)',NA,lst[[stock_name]]$priceg)
    #Have to convert back because gsubbed
    lst[[stock_name]]$priceg<-as.numeric(lst[[stock_name]]$priceg)
}

#Combine all data.frames in list
udf <- ldply(lst, data.frame)

#Join with set_return on i
set_return<-read.csv('set_return.csv')
udf<-join(udf,set_return)

#Save to udf.csv
write.csv(udf,file='udf.csv',row.names=FALSE)
```

# Variable Description
The dataset includes annual fundamental data of 463 Thai stocks from 2002-2015. The variables are divided into 7 groups same as the financial statement format of [Morningstar](http://www.morningstar.com) namely:

* ```Income Statement```: column 4-31
* ```Balance Sheet```: column 32-81
* ```Cash Flow Statement```: column 82-115
* ```Valuation Ratios```: column  116-127 and 188
* ```Valuation and Quality Indicators```: column 128-154
* ```Common Size Ratios```: column 155-174
* ```Per Share Data```: column 175-185

Other variables include:

* ```.id```, ```symbol```: stock symbols (column 1,2)
* ```i```: year indicator (column 3)
* ```fwd_price```: forward price (column 186)
* ```priceg```: forward growth (%) (column 187). At year t, priceg = (price year t+1) / (price at year t)
* ```set_return```: annual SET TRI return (%) (column 189)

The list of variables in ```udf.csv``` is as follows. Note that some variables appear in several groups. For instance, ```monthendstockprice``` and ```monthendstockprice.1```.

```{r}
udf<-read.csv('udf.csv')
colnames(udf)
```
